results

### 2020-10-11 analysis/results.Rmd
re-render results

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(ggeconodist)
library(gt)
```

@@

```{r}
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
```

```{r}
# paratext pattern
para <- c('^Author Index$|^Back Cover|^Contents$|^Contents:|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page')
```

```{r}
hybrid_md <- readr::read_csv(here::here("data", "hybrid_md.csv"))
els_tdm <- readr::read_csv(here::here("data", "els_tmd_links.csv"))
invoice_df <-  readr::read_csv(here::here("data", "invoice_df.csv"))

hybrid_df <- els_tdm %>% 
  select(doi, URL) %>% 
  inner_join(invoice_df, by = c("URL" = "tdm_url")) %>% 
  select(-URL) %>% 
  inner_join(hybrid_md, by = c("doi")) %>% 
  distinct() %>%
  filter(!grepl(para, title))

flipped_jns_df <- els_jns %>%
  filter(issn %in% flipped_jns$issn) %>%
  select(oa_model, issn, year)

hybrid_df <- left_join(hybrid_df, flipped_jns_df, by = c("issn", "issued_year" = "year")) %>%
  filter(!oa_model %in% "Open Access") %>%
  select(-oa_model)
```

@@

```{r}
eligible_jns <- hybrid_df %>% 
  filter(oa_archive == FALSE) %>% 
  distinct(issn)
```

```{r}
# required data spreadsheets
# pub volume
els_pub_volume <- readr::read_csv(
  here::here("data", "article_volume_para.csv")
  ) 
# eligble journals 
eligible_jns <- hybrid_df %>% 
  filter(oa_archive == FALSE) %>% 
  distinct(issn)
# pub volume eligble data
hybrid_volume <- els_pub_volume %>%
  filter(issn %in% eligible_jns$issn) %>%
  inner_join(hybrid_df, by = c("issn", "issued_year")) %>%
  filter(oa_archive == FALSE) %>%
  filter(!grepl("^S", page))
# backup 
write_csv(hybrid_volume, here::here("data", "hybrid_volume_eligible.csv"))

els_articles <- 
  hybrid_volume %>%
  distinct(issued_year, issn, articles) %>%
  group_by(issued_year) %>%
  summarise(
    articles_n = sum(articles),
            articles_mean = mean(articles),
            articles_sd = sd(articles)
    )
els_yr <- 
  hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(journals = n_distinct(issn)) %>%
  inner_join(els_articles, by = "issued_year")

# oa
oa_per_jn_year <- hybrid_volume %>%
  group_by(issued_year, issn) %>%
  summarise(oa = n(), .groups = "drop")
oa_ind <- oa_per_jn_year %>%
  group_by(issued_year) %>%
  summarise(oa_n = sum(oa),
            oa_mean = mean(oa),
            oa_sd = sd(oa)) 
# oa uptake 
oa_uptake_ind <- hybrid_volume %>% 
  distinct(issued_year, issn, articles) %>%
  inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# oa uptake all 
oa_uptake_15_19 <- hybrid_volume %>% 
  distinct(issued_year, issn, articles) %>%
  inner_join(oa_per_jn_year, by = c("issn", "issued_year")) %>%
    group_by(issn) %>%
    summarise(oa = sum(oa),
              articles = sum(articles)) %>%
    mutate(prop = oa / articles)

# all 
total_df <- tibble::tibble(
  issued_year = "2015-19",
  journals = hybrid_volume %>% distinct(issn) %>% nrow(),
  articles_n = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sum(),
  articles_mean = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% mean(),
  articles_sd = hybrid_volume %>% distinct(issn, articles, issued_year) %>% pull(articles) %>% sd(),
  oa_n = nrow(hybrid_volume),
  oa_mean = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% mean(),
  oa_sd = hybrid_volume %>% group_by(issn) %>% summarise(oa = n()) %>% pull(oa) %>% sd(),
  oa_uptake = oa_n / articles_n,
  oa_uptake_mean = mean(oa_uptake_15_19$prop),
  oa_uptake_sd = sd(oa_uptake_15_19$prop)
         )

ind_df <- inner_join(els_yr, oa_ind, by = "issued_year") %>%
  mutate(oa_uptake = oa_n / articles_n) %>%
  inner_join(oa_uptake_ind, by = "issued_year") %>%
  mutate(issued_year = as.character(issued_year)) %>%
  bind_rows(total_df) %>%
  mutate(oa_uptake = oa_uptake * 100) %>%
  mutate(oa_uptake_mean = oa_uptake_mean * 100) %>%
  mutate(oa_uptake_sd = oa_uptake_sd * 100)

ind_summary_table <- ind_df %>% 
  pivot_longer(journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model. Of those, `r hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio. These journals published a total of `r total_df$oa_n` immediate open access articles between 2015 and 2019, corresponding to a share of `r total_df$oa_uptake * 100`%. Table present the high-level findings by year, illustrating a moderate growth in hybrid open access. Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_df %>% filter(issued_year == "2015") %>% pull(oa_n)` in 2015 to `r ind_df %>% filter(issued_year == "2019") %>% pull(oa_n)` in 2019. However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_df %>% filter(issued_year == 2015) %>% pull(oa_uptake)`% to `r ind_df %>% filter(issued_year == 2019) %>% pull(oa_uptake)`%, because the number of articles published also increased during this five-years period.

```{r}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~ind_name,
  "Elsevier Hybrid Journals with OA article(s)", "journals", "Journals",
  "Articles published", "articles_n", "Total",
  "Articles published", "articles_mean", "Avg.",
  "Articles published", "articles_sd", "SD",
  "OA Articles published", "oa_n", "Total OA",
  "OA Articles published", "oa_mean", "Avg.",
  "OA Articles published", "oa_sd", "SD",
  "OA Uptake (in%)", "oa_uptake", "Percentage",
  "OA Uptake (in%)", "oa_uptake_mean", "Avg.",
  "OA Uptake (in%)", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind) %>%
  gt(rowname_col = "ind_name", groupname_col = "ind_group") %>%
  fmt_number(contains("20"), decimals = 1) %>%
  fmt_number(contains("20"), 
             rows = c("Total", "Total OA", "Journals"),
             decimals = 0) %>%
  tab_style(
    style = cell_fill(
      color = "grey90"
    ),
    locations = list(
      cells_body(
        columns = "2015-19"
      ),
      cells_column_labels(
        columns = "2015-19"
      )
    )
  ) %>%
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )
  )  %>%
   tab_style(
    style = cell_text(color = "black", style = "italic"),
    locations = list(
      cells_row_groups())
    ) %>%
  cols_width(
    2 ~ px(125),
    3:7 ~ px(80),
    8 ~ px(110)
  ) %>%
  tab_options(
    row_group.border.top.width = px(3),
    row_group.border.top.color = "white",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "white",
    table.border.top.width = px(3),
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3),
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(2),
    stub.border.color = "white"
  )
```

```{r}
els_jn_pub_volume <- els_pub_volume %>%
  filter(issn %in% eligible_jns$issn) %>%
  group_by(issued_year, issn) %>%
  summarise(articles = sum(articles),
            journals = n_distinct(issn))
hybrid_oa_per_jn <- hybrid_volume %>%
  group_by(issued_year, issn) %>%
  summarise(oa_articles = n())%>%
  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles) 
```

```{r}
perc_df <- hybrid_oa_per_jn %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

@@

```{r, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 18%."}
ggplot(hybrid_oa_per_jn, aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.18)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Years",
       y = "Open Access Uptake")
# licenses used
```

```{r}
# per year
yrl_volume <- hybrid_oa_per_jn %>% 
  group_by(issued_year) %>% 
  summarise(articles = sum(articles))

hybrid_license <- hybrid_volume %>%
  filter(oa_archive == FALSE) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
  group_by(cc_norm, issued_year) %>%
  summarise(cc_articles = n())
```

@@

Figure presents a breakdown by Creative Commons license variant used for hybrid open access by year. Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals. `r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes. This proportion remained the same between 2015-19.

```{r, fig.cap = "Immediate open access articles in hybrid journals by open content license and year. The blue area represent articles permitting commercial adaption under a Creative Commons CC-BY license, while the purple area displays articles under a less permissive Commons CC-BY-ND license."}
hybrid_license %>%
  ggplot(aes(factor(issued_year), cc_articles, fill = cc_norm, group = cc_norm)) +
  geom_area(stat = "identity",
            color = "white",
            alpha = 0.8,
            position = position_stack(reverse = T)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = function(x)
        format(x, big.mark = ",", scientific = FALSE),
      breaks = scales::pretty_breaks()
  ) +
  scale_fill_manual(values = 
                      c(`cc-by` = "#0093c7",
  `cc-by-nc` = "#B52141")) +
  labs(x = "Publication year", y = "Total OA articles") +
  theme_minimal_hgrid() +
  theme(legend.position = "top",
        legend.justification = "right") +
  guides(fill = guide_legend("Creative Commons license variant"))
```

@@

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_volume_eligible.csv"))

@@

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  rename(yrl_volume = articles) %>%
  left_join(tt, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc"))

@@

```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)'}
sponsor_country %>%
  mutate(country_factor = fct_reorder(country_factor, n)) %>%
  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
) %>%
  ggplot(aes(country_factor, n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  coord_flip() +
  theme_minimal_vgrid() +
  labs(y = "Number of invoiced OA articles", x = NULL)
```

@@

```{r}
hybrid_invoice_tt <- hybrid_invoice %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
 group_by(oa_sponsor) %>%
  summarise(journals = n_distinct(issn),
            articles = n_distinct(doi)) %>%
  mutate(prop = articles / sum(articles))
hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
  ungroup() %>%
  filter(cc_norm == "cc-by") %>%
  group_by(oa_sponsor) %>%
  summarise(cc_by = n_distinct(doi)) 
hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
  filter(in_open_apc == TRUE) %>%
  group_by(oa_sponsor) %>%
  summarise(in_open_apc =  n_distinct(doi))
ind_sponsor_df <- left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
  mutate(cc_by = cc_by / articles,
         in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
         in_open_apc = in_open_apc / articles) %>% 
  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
# overall
overall_ind_sponsor_df <- tibble(oa_sponsor = "All", journals = length(unique(hybrid_invoice$issn)), articles = length(unique(hybrid_invoice$doi)), prop = 1, cc_by = filter(hybrid_invoice, cc_norm == "cc-by") %>% nrow() / length(unique(hybrid_invoice$doi)), in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)))

>>>>

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
```

@@

@@

```{r summary_statistics_per_year}
hybrid_oa_volume <- readr::read_csv(here::here("data", "hybrid_oa_volume.csv"))
# journal and article metrics **per year**
hybrid_oa_per_year <- hybrid_oa_volume %>%
  group_by(issued_year) %>%
  # distinct jns with at least one article per year
  filter(oa > 0) %>%
  summarise(
      # distinct jns with at least one article per year
    n_journals = n(),
    # article metrics
    articles_n = sum(articles),
    articles_mean = mean(articles),
    articles_sd = sd(articles),
    # oa articles metrics
    oa_n = sum(oa),
    oa_mean = mean(oa),
    oa_sd = sd(oa)
    ) %>%
  mutate(oa_uptake = oa_n / articles_n)
# oa uptake per year
oa_uptake_per_year <- hybrid_oa_volume %>% 
  filter(oa > 0) %>%
  mutate(prop = oa / articles) %>%
  group_by(issued_year) %>%
  summarise(oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop))
# merge indicators
ind_per_year <- inner_join(hybrid_oa_per_year, oa_uptake_per_year, by = "issued_year")
```

```{r summary_statistics_2015_19}
# five-years article volume
all_jn_articles <- hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles)
    ) %>%
  ungroup() %>%
  summarise(
    n_journals = n(),
    articles_n = sum(articles_p_jn),
    articles_mean = mean(articles_p_jn),
    articles_sd = sd(articles_p_jn)
  )
# five-years OA article volume
all_oa_articles <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    oa_p_jn = sum(oa)
    ) %>%
  ungroup() %>%
  summarise(
    oa_n = sum(oa_p_jn),
    oa_mean = mean(oa_p_jn),
    oa_sd = sd(oa_p_jn)
  )
# five-years OA uptake rate
oa_uptake <- 
  hybrid_oa_volume %>%
  group_by(issn) %>%
  summarise(
    # articles
    articles_p_jn = sum(articles),
    oa_p_jn = sum(oa)
    ) %>%
  mutate(prop = oa_p_jn / articles_p_jn) %>%
  ungroup() %>%
  summarise(oa_uptake = sum(oa_p_jn) / sum(articles_p_jn),
            oa_uptake_mean = mean(prop),
            oa_uptake_sd = sd(prop)
  )
# bringing it all together
five_years_ind <- bind_cols(issued_year = "2015-19", all_jn_articles, all_oa_articles, oa_uptake)

ind_summary_table <- ind_per_year %>%
  mutate(issued_year = as.character(issued_year)) %>% 
  bind_rows(five_years_ind) %>%
  mutate(across(contains("uptake")) * 100) %>%
  pivot_longer(n_journals:oa_uptake_sd, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count)
```

On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model. Of those, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_oa_volume %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio. These journals published a total of `r sum(hybrid_oa_volume$oa)` immediate open access articles between 2015 and 2019, corresponding to a share of `r round(sum(hybrid_oa_volume$oa) / sum(hybrid_oa_volume$articles) * 100, 1)`%. Table present the high-level findings by year, illustrating a moderate growth in hybrid open access. Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r hybrid_oa_volume %>% filter(issued_year == "2015") %>% pull(oa) %>% sum()` in 2015 to `r hybrid_oa_volume %>% filter(issued_year == "2019") %>% pull(oa) %>% sum()` in 2019. However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r round(ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake) * 100, 1)`% to `r round(ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake) *100, 1)`%, because the number of articles published also increased during this five-years period.

```{r}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~ind_name,
  "Elsevier Hybrid Journals with OA article(s)", "n_journals", "Journals",
  "Articles published", "articles_n", "Total",
  "Articles published", "articles_mean", "Avg.",
  "Articles published", "articles_sd", "SD",
  "OA Articles published", "oa_n", "Total OA",
  "OA Articles published", "oa_mean", "Avg.",
  "OA Articles published", "oa_sd", "SD",
  "OA Uptake (in%)", "oa_uptake", "Percentage",
  "OA Uptake (in%)", "oa_uptake_mean", "Avg.",
  "OA Uptake (in%)", "oa_uptake_sd", "SD"
)
gt_1 <- inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind) %>%
  gt(rowname_col = "ind_name", groupname_col = "ind_group") %>%
  fmt_number(contains("20"), decimals = 1) %>%
  fmt_number(contains("20"), 
             rows = c("Total", "Total OA", "Journals"),
             decimals = 0) %>%
  tab_style(
    style = cell_fill(
      color = "grey90"
    ),
    locations = list(
      cells_body(
        columns = "2015-19"
      ),
      cells_column_labels(
        columns = "2015-19"
      )
    )
  ) %>%
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )
  )  %>%
   tab_style(
    style = cell_text(color = "black", style = "italic"),
    locations = list(
      cells_row_groups())
    ) %>%
  cols_width(
    2 ~ px(125),
    3:7 ~ px(80),
    8 ~ px(110)
  ) %>%
  tab_options(
    row_group.border.top.width = px(3),
    row_group.border.top.color = "white",
    row_group.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.top.color = "white",
    table.border.top.width = px(3),
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3),
    column_labels.border.bottom.color = "black",
    column_labels.border.bottom.width = px(2),
    stub.border.color = "white"
  )
gt_1
```

```{r}
perc_df <- hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
  group_by(issued_year) %>% 
  summarize(median = median(prop),
            p_10 = quantile(prop, 0.1),
            p_95 = quantile(prop, 0.95),
            iqr = quantile(prop, 3/4) - quantile(prop, 1/4)) %>%
  ungroup()
```

@@

```{r, fig.cap = "Open access uptake per Elsevier hybrid journal in percent by year, visualized as box-and-whiskers plot. The Y-axis is limited to an open access share of 18%."}
hybrid_oa_volume %>%
  mutate(prop = oa / articles) %>%
  filter(oa > 0) %>%
ggplot(aes(factor(issued_year), prop)) + 
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + 
  coord_cartesian(ylim=c(0,0.2)) +
  theme_minimal_hgrid() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L),
      expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Publication Years",
       y = "Open Access Uptake")
```

```{r}
# per year
hybrid_articles <- readr::read_csv(here::here("data", "hybrid_articles.csv"))

hybrid_license <- hybrid_articles %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
  group_by(cc_norm, issued_year) %>%
  summarise(cc_articles = n())
```

@@

Figure presents a breakdown by Creative Commons license variant used for hybrid open access by year. Elsevier provided commercial (cc-by) and a less permissive non-commercial license (cc-by-nc) for open access publishing in otherwise subscription-based journals. `r  round(hybrid_license %>% group_by(cc_norm) %>% summarise(n = sum(cc_articles)) %>% mutate(prop = n / sum(n)) %>% filter(cc_norm == "cc-by-nc-nd") %>% pull(prop) * 100,0)` percent of open access articles can be re-used for non-commercial purposes. This proportion remained the same between 2015-19.

```{r, fig.cap = "Immediate open access articles in hybrid journals by open content license and year. The blue area represent articles permitting commercial adaption under a Creative Commons CC-BY license, while the purple area displays articles under a less permissive Commons CC-BY-ND-NC license."}
hybrid_license %>%
  ggplot(aes(factor(issued_year), cc_articles, fill = cc_norm, group = cc_norm)) +
  geom_area(stat = "identity",
            color = "white",
            alpha = 0.8,
            position = position_stack(reverse = T)) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = function(x)
        format(x, big.mark = ",", scientific = FALSE),
      breaks = scales::pretty_breaks()
  ) +
  scale_fill_manual(values = 
                      c(`cc-by` = "#B52141",
                        `cc-by-nc-nd` = "#0093c7")) +
  labs(x = "Publication year", y = "Total OA articles") +
  theme_minimal_hgrid() +
  theme(legend.position = "top",
        legend.justification = "right") +
  guides(fill = guide_legend("Creative Commons license variant"))
```

@@

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv"))

@@

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
 # rename(yrl_volume = articles) %>%
  left_join(tt, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))

@@

```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)'}
sponsor_country %>%
  mutate(country_factor = fct_reorder(country_factor, n)) %>%
  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
) %>%
  ggplot(aes(country_factor, n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::number_format(big.mark = " ")) +
  coord_flip() +
  theme_minimal_vgrid() +
  labs(y = "Number of invoiced OA articles", x = NULL)
```

@@

```{r}
hybrid_invoice_tt <- hybrid_invoice %>%
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
hybrid_invoice_agg_general <- hybrid_invoice_tt %>%
 group_by(oa_sponsor) %>%
  summarise(journals = n_distinct(issn),
            articles = n_distinct(doi)) %>%
  mutate(prop = articles / sum(articles))
hybrid_invoice_agg_cc <- hybrid_invoice_tt %>%
  ungroup() %>%
  filter(cc_norm == "cc-by") %>%
  group_by(oa_sponsor) %>%
  summarise(cc_by = n_distinct(doi)) 
hybrid_invoice_agg_oapc <- hybrid_invoice_tt %>%
  filter(in_open_apc == TRUE) %>%
  group_by(oa_sponsor) %>%
  summarise(in_open_apc =  n_distinct(doi))
ind_sponsor_df <- left_join(hybrid_invoice_agg_general, hybrid_invoice_agg_cc, by = "oa_sponsor") %>%
  left_join(hybrid_invoice_agg_oapc, by = "oa_sponsor") %>%
  mutate(cc_by = cc_by / articles,
         in_open_apc = ifelse(is.na(in_open_apc), 0, in_open_apc),
         in_open_apc = in_open_apc / articles) %>% 
  mutate(oa_sponsor = fct_reorder(oa_sponsor, articles)) %>%
  mutate(oa_sponsor = forcats::fct_relevel(oa_sponsor, "Other", after = 0L)) %>% arrange(desc(oa_sponsor))
# overall
overall_ind_sponsor_df <- tibble(oa_sponsor = "All", journals = length(unique(hybrid_invoice$issn)), articles = length(unique(hybrid_invoice$doi)), prop = 1, cc_by = filter(hybrid_invoice, cc_norm == "cc-by") %>% nrow() / length(unique(hybrid_invoice$doi)), in_open_apc = filter(hybrid_invoice, in_open_apc == TRUE) %>% nrow() / length(unique(hybrid_invoice$doi)))
